---
title: "Experimental design & metadata"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
 rmarkdown::html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
    highlight: haddock
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# Setup

## Libraries

```{r}
library("dplyr")
#install.packages("lubridate")
library("lubridate")
library("ggplot2")
library("cowplot")
#install.packages("Rmisc")
library("Rmisc")
#install.packages("coxme")
library("coxme")
library("car")
#install.packages("bbmle")
library("bbmle")
#install.packages("emmeans")
library("emmeans")
#install.packages("DHARMa")
library("DHARMa")
#install.packages("multcomp")
library("multcomp")
#install.packages("multcompView")
library("multcompView")
library("glmmTMB")
```

## Read in data

```{r}
setwd("/Volumes/Macintosh HD/Users/nicolakriefall/nicolagk@hawaii.edu - Google Drive/My Drive/Mosquito_business/Mosquito_microbes/01.data_expdesign")

##metadata
##same file that goes into phyloseq
metadata <- read.csv("../02.process_asvs/counts_taxa_info/metadata_cleaned.csv")
row.names(metadata) <- metadata$sample_name

metadata2 <- metadata %>% 
  mutate(newday=ifelse(
    day<=12, "early",
    ifelse(
      day>12 & day<=20, "mid",
      ifelse(
        day>20,"late","NA"
      ))))

metadata2 %>%
  subset(type=="A.albopictus") %>%
  group_by(newday) %>%
  dplyr::count()
```

# Experimental info

## Variable notes

Abbreviations

- Infusion
  - **OL** = 'ohi'a lehua
  - **SG** = strawberry guava 
  - **SW** = sterile water
- Dispersal
  - **N** = 'no' dispersal (filtered)
  - **D** = dispersal (unfiltered)
- Temperature
  - **H** = hot
  - **C** = cool

# Date things & plot

Dates: Aug. 15th - Sep 17th and 22nd, 2019

## Setup things

```{r}
metadata2$date_cleaned <- mdy(metadata2$date.collected)

##checking that experimental day & date line up:
# ggplot(metadata2,aes(x=date_cleaned,y=day))+
#  geom_point()

meta.mq <- subset(metadata2,stage=="Adult")

meta.mq$inf_temp <- paste0(meta.mq$infusion,"_",meta.mq$temperature)
```

## Experimental timeline plot (Fig. 1 panel)

```{r}
meta.mq$fake <- rep(30,nrow(meta.mq))
#adding some fake points to get the shapes into the legend via geom_point

meta.mq$temperature <- sub("C","Cool",meta.mq$temperature)
meta.mq$temperature <- sub("H","Warm",meta.mq$temperature)

ggplot(meta.mq,aes(x=day,fill=inf_temp))+
  geom_point(aes(shape=sex,y=fake))+
  scale_shape_manual(values=c(21,23),name="Events",labels=c("Dispersal","Disp./water sample"))+
  geom_histogram(position=position_dodge(),color="black")+
  theme_cowplot()+
  scale_fill_manual(values=c("#B5D69F","#7DBA54","#C488A2","#992559","#B6F6EE","#60eede"),name="Infusion, temp.",labels=c("OL_cool","OL_warm","SG_cool","SG_warm","PW_cool","PW_warm"))+
  #ggtitle("Time to emergence")+
  xlab("Experimental day")+
  ylab("Adults emerged")+
  coord_cartesian(xlim = c(2,40))+
  #ylim(-1,28)+
  ylim(-1,19)+
  #scale_alpha_manual(values=c(0.65,1),name="Thermal regime",labels=c("Cool","Warm"))+
  annotate("point", x=c(2,6,8,10,14,16,18), y=c(rep(-1,7)), shape=21, size=1.5,fill="white")+
  annotate("point", x=c(4,12,20), y=c(-1,-1,-1), shape=23, size=2,fill="white")+
  facet_grid(temperature~.)+
  theme(text=element_text(size=12),axis.text.x=element_text(size=12),axis.text.y=element_text(size=12))

#ggsave(file="exp1_timelineNov23.pdf",height=3.25,width=5.25)
```

## Average time to emergence (Supp. fig)

```{r}
meta.mq.se <- summarySE(meta.mq,measurevar="day",groupvars=c("inf_temp","infusion"))

ggplot(meta.mq,aes(x=infusion,y=day,fill=inf_temp,shape=inf_temp))+
  geom_jitter(alpha=0.5,position=position_jitterdodge(jitter.width=0.3),color="gray")+
  geom_errorbar(data=meta.mq.se,aes(ymax=day+se,ymin=day-se),width=0.2,color="black",position=position_dodge(width=0.6))+
  geom_point(data=meta.mq.se,size=2.5,position=position_dodge(width=0.6))+
  theme_cowplot()+
  xlab("Infusion")+
  ylab("Day of emergence")+
  scale_fill_manual(values=c("#B5D69F","#7DBA54","#C488A2","#992559","#B6F6EE","#60eede"),name="Infusion, temp.",labels=c("OL_cool","OL_warm","SG_cool","SG_warm","PW_cool","PW_warm"))+
  scale_shape_manual(values=c(22,24,22,24,22,24),name="Infusion, temp.",labels=c("OL_cool","OL_warm","SG_cool","SG_warm","PW_cool","PW_warm"))+
  scale_x_discrete(labels=c("OL","SG","PW"))
#ggsave("plot.emerge.rate.pdf",width=5,height=3.5)
```

### Time to emergence stats

```{r}
##summary stats
summarySE(meta.mq,measurevar="day",groupvars="infusion")
summarySE(meta.mq,measurevar="day",groupvars=c("infusion","temperature"))

##coxme
meta.mq$emergence <- 1

cox.all <- coxme(Surv(day, emergence)~infusion+temperature+dispersal+(1|mesocosm), data=meta.mq)
summary(cox.all)

Anova(cox.all)

# Response: Surv(day, emergence)
#             Df    Chisq Pr(>Chisq)    
# infusion     2 100.6853    < 2e-16 ***
# temperature  1  95.8436    < 2e-16 ***
# dispersal    1   3.6092    0.05746 .  

cox.all.int <- coxme(Surv(day, emergence)~infusion*temperature*dispersal+(1|mesocosm), data=meta.mq)
Anova(cox.all.int)

cox.all.int1 <- coxme(Surv(day, emergence)~infusion*temperature+dispersal+(1|mesocosm), data=meta.mq)

cox.all.int2 <- coxme(Surv(day, emergence)~infusion+temperature*dispersal+(1|mesocosm), data=meta.mq)

cox.all.int3 <- coxme(Surv(day, emergence)~infusion*dispersal+temperature+(1|mesocosm), data=meta.mq)

AICtab(cox.all,cox.all.int,cox.all.int1,cox.all.int2,cox.all.int3)
#cox.all best

##infusion
cox.all.noinf <- coxme(Surv(day, emergence)~temperature+dispersal+(1|mesocosm), data=meta.mq)
anova(cox.all,cox.all.noinf) #p***
##temperature
cox.all.notem <- coxme(Surv(day, emergence)~infusion+dispersal+(1|mesocosm), data=meta.mq)
anova(cox.all,cox.all.notem) #p***
##dispersal
cox.all.nodis <- coxme(Surv(day, emergence)~infusion+temperature+(1|mesocosm), data=meta.mq)
anova(cox.all,cox.all.nodis) #0.06 .

##groups
cox.all.em <- emmeans(cox.all,~infusion+temperature)
multcomp::cld(cox.all.em)
```

## Step plot by date

```{r}
# surv.mq <-
#   meta.mq %>%
#   mutate(
#     status = recode(stage, "Adult" = 1)
#   )

surv.mq2 <- meta.mq %>%
  group_by(infusion,day) %>%
  dplyr::summarize(num=n()) %>%
  dplyr::mutate(count.inf=cumsum(num)) %>%
  dplyr::mutate(prop=(count.inf/240)*100)

# ggplot(surv.mq2,aes(x=day,y=prop,color=infusion))+
#   geom_step(linewidth=1)+
#   #xlim(0,39)+
#   theme_linedraw()+
#   xlab("Experimental day")+
#   ylab("Percent of larvae hatched")+
#   scale_color_manual(values=c("#7DBA54","#992559","#60eede"),name="Infusion")

surv.mq3 <- surv.mq2 %>%
  as.data.frame() %>%
  add_row(infusion="OL",day=0,num=0,count.inf=0,prop=0) %>%
  add_row(infusion="SG",day=0,num=0,count.inf=0,prop=0) %>%
  add_row(infusion="SW",day=0,num=0,count.inf=0,prop=0) %>%
  add_row(infusion="OL",day=39,num=NA,count.inf=NA,prop=42.9166667) %>%
  add_row(infusion="SG",day=39,num=NA,count.inf=NA,prop=18.7500000) %>%
  add_row(infusion="SW",day=39,num=NA,count.inf=NA,prop=20.4166667)

ggplot(surv.mq3,aes(x=day,y=prop,color=infusion))+
  geom_step(linewidth=1)+
  #xlim(0,39)+
  theme_linedraw()+
  xlab("Experimental day")+
  ylab("Percent of larvae hatched")+
  scale_color_manual(values=c("#7DBA54","#992559","#60eede"),name="Infusion")

## survival plot stuff I didn't end up using
# install.packages(c("survival", "lubridate", "ggsurvfit", "gtsummary", "tidycmprsk"))
# remotes::install_github("zabore/condsurv")
# remotes::install_github("zabore/ezfun")
# library(gtsummary)
# library(tidycmprsk)
# library(condsurv)

# library(survival)
# #install.packages("ggsurvfit")
# library(ggsurvfit)
# 

# 
# Surv(surv.mq$day, surv.mq$status)[1:10]
# s1 <- survfit(Surv(day,status) ~ 1, data = surv.mq)
# str(s1)
# 
# survfit2(Surv(day,status) ~ 1, data = surv.mq) %>% 
#   ggsurvfit() +
#   labs(
#     x = "Days",
#     y = "Overall survival probability"
#   )+ 
#   add_confidence_interval()
```

## Emergence stats

### Setup

```{r}
#subset mosquito adults from the data
main.mq <- subset(metadata2, type=="A.albopictus")

#summarize how many per mesocosm
surv <- as.data.frame(summary(as.factor(main.mq$mesocosm)), optional=TRUE)
##there should be 72 mesocosms here..
surv$mesocosm <- rownames(surv)
colnames(surv) <- c("count","mesocosm")

##selecting mesocosm metadata
meta.notspecial <- subset(metadata2,type=="Microbial Water")

main.meso <- meta.notspecial %>%
  dplyr::select(mesocosm, infusion, temperature, dispersal)
##keeping only unique data
main.meso.uniq <- unique(main.meso)
##merging with survival data
surv2 <- merge(surv,main.meso.uniq,by="mesocosm",all=T)
surv2[is.na(surv2)] <- 0
```

### Success per mesocosm plot (Fig. 1 panel)

```{r}
surv2$inf.temp <- paste0(surv2$infusion,surv2$temperature)

surv2$count10 <- (surv2$count)*10
#meta.mq.se <- summarySE(meta.mq,measurevar="day",groupvars=c("inf_temp","infusion"))

surv2.se <- summarySE(surv2,measurevar="count10",groupvars=c("inf.temp","infusion"))
surv2.se

ggplot(surv2.se,aes(x=infusion,y=count10,fill=inf.temp))+
  geom_bar(stat="identity",position=position_dodge(),color="black")+
  scale_fill_manual(values=c("#B5D69F","#7DBA54","#C488A2","#992559","#B6F6EE","#60eede"),name="Infusion, temp.",labels=c("OL, cool","OL, warm","SG, cool","SG, warm","PW, cool","PW, warm"))+
  theme_cowplot()+
  geom_errorbar(aes(ymin=count10-se,ymax=count10+se,width=0.2),position=position_dodge(1))+
  ylim(0,55)+
  theme()+
  ylab("Emerged/meso. (%)")+
  xlab("")+
  scale_x_discrete(labels=c("OL","SG","PW"))
  #geom_text(aes(label=round(count10)),vjust=-0.25,position=position_dodge(1))

#ggsave("gg.prop.pdf",width=4,height=2.5)
```

### Stats for success

```{r}
# mod.surv <- glmmTMB(cbind(count,10-count)~infusion+temperature+dispersal+(1|mesocosm), family=binomial(link="logit"), data=surv2)
# mod.surv
mod.surv <- glmmTMB(cbind(count,10-count)~infusion+temperature+dispersal+(1|mesocosm), family="binomial", data=surv2)
#mod.surv

mod.surv.int <- glmmTMB(cbind(count,10-count)~infusion*temperature*dispersal+(1|mesocosm), family="binomial", data=surv2)

mod.surv.int1 <- glmmTMB(cbind(count,10-count)~infusion+temperature*dispersal+(1|mesocosm), family="binomial", data=surv2)

mod.surv.int2 <- glmmTMB(cbind(count,10-count)~infusion*temperature+dispersal+(1|mesocosm), family="binomial", data=surv2)

mod.surv.int3 <- glmmTMB(cbind(count,10-count)~infusion*dispersal+temperature+(1|mesocosm), family="binomial", data=surv2)

AICtab(mod.surv,mod.surv.int,mod.surv.int1,mod.surv.int2,mod.surv.int3)
#interaction better

Anova(mod.surv.int2)

anova(mod.surv,mod.surv.int2) #p<0.05*

mod.surv.int2.nodis <- glmmTMB(cbind(count,10-count)~infusion*temperature+(1|mesocosm), family="binomial", data=surv2)

anova(mod.surv.int2,mod.surv.int2.nodis) #0.53

mod.surv.int2.resid <- simulateResiduals(fittedModel = mod.surv.int2, plot = T)

mod.surv.int2.em <- emmeans(mod.surv.int2,~infusion*temperature)
multcomp::cld(mod.surv.int2.em)
```

# qPCR things

## Setup

```{r}
meta.mq$wolb <- ((1/meta.mq$Wolb_ct1)+(1/meta.mq$Wolb_ct2))/(1/meta.mq$Act_ct)
hist(meta.mq$wolb)

ggplot(meta.mq,aes(x=infusion,y=wolb,color=temperature))+
  geom_boxplot()+
  facet_wrap(~sex)
```

## Stats

```{r}
#Wolbachia
hist(meta.mq$wolb)
hist(log(meta.mq$wolb))

##full model
wolb.full <- glmmTMB(log(wolb)~dispersal+newday+infusion+temperature+sex+(1|mesocosm), data=meta.mq)
##no dispersal
wolb.nodis <- glmmTMB(log(wolb)~newday+infusion+temperature+sex+(1|mesocosm), data=meta.mq)
##no time
wolb.noday <- glmmTMB(log(wolb)~dispersal+infusion+temperature+sex+(1|mesocosm), data=meta.mq)
##no infusion
wolb.noinf <- glmmTMB(log(wolb)~dispersal+newday+temperature+sex+(1|mesocosm), data=meta.mq)
##no temperature
wolb.notem <- glmmTMB(log(wolb)~dispersal+newday+infusion+sex+(1|mesocosm), data=meta.mq)
##no sex
wolb.nosex <- glmmTMB(log(wolb)~dispersal+newday+infusion+temperature+(1|mesocosm), data=meta.mq)

AICtab(wolb.full,wolb.nodis,wolb.noday,wolb.noinf,wolb.notem,wolb.nosex)

anova(wolb.full,wolb.nodis) #ns
anova(wolb.full,wolb.noday) #ns
anova(wolb.full,wolb.noinf) #ns
anova(wolb.full,wolb.notem) #ns
anova(wolb.full,wolb.nosex) #0.07 . 

shapiro.test(residuals(wolb.full)) 
qqnorm(residuals(wolb.full))
qqline(residuals(wolb.full), col="red")

wolb.full.resid <- simulateResiduals(fittedModel = wolb.full, plot = T)
```

