---
title: "Alpha diversity"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
 rmarkdown::html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
    highlight: haddock
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# Setup

## Libs

```{r}
library(car) 
library("Rmisc")
library("ggpubr")
library("cowplot")
library("lme4")
library("phyloseq")
library("glmmTMB")
library("emmeans")
library(multcomp)
#install.packages("multcompView")
library(multcompView)
library("bbmle")
library("DHARMa")
library("nlme")
library("bestNormalize")
```

## Data

```{r}
setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Mosquito_business/Mosquito_microbes/04.microbiome_analysis/04a.diversity")

ps.clean <- readRDS("../../02.process_asvs/ps.clean.trim.less.rds")
ps.clean 
#ps.clean <- readRDS("../../02.process_asvs/ps.clean.trim.rare9200.rds")
#ps.clean

samdf <- data.frame(ps.clean@sam_data)

##made below
#df.div <- read.csv("div.trim.csv",row.names=1)
```

# Alpha diversity calculations (first run through)

```{r}
##option to run rarefied stuff:
#ps.clean <- ps.rare

df.all <- data.frame(estimate_richness(ps.clean, split=TRUE, measures=c("Shannon","InvSimpson","Observed")))

df.all$sample_name <- rownames(df.all)

df.all.div <- merge(df.all,samdf,by="sample_name") #add sample data

#shannon diversity divided by species richness
df.all.div$even <- df.all.div$Shannon/(log(df.all.div$Observed))

df.div <- df.all.div

#adding most recent sample counts
sums <- data.frame(sample_sums(ps.clean))
colnames(sums) <- c("lib_size_trim")
sums$sample_name <- row.names(sums)

df.div <- merge(df.div,sums,by="sample_name")

#write.csv(df.div,"div.trim.csv")
```

# Analyzing alpha metrics

## All data

Very strange/interesting pattern where the water types that shouldn't have a lot of diversity have a lot of OTUs, but not a lot of reads... 

```{r}
ggplot(df.div,aes(x=type,y=Observed,color=infusion))+
  geom_boxplot()+
  geom_jitter()

ggplot(df.div,aes(x=type,y=lib_size_trim,color=infusion))+
  geom_boxplot()+
  geom_jitter(position=position_jitterdodge())
```

## Plotz

```{r}
#just experimental types
df.div.exp <- subset(df.div,type=="Microbial Water"|type=="A.albopictus")

df.div.exp$inf.temp <- paste0(df.div.exp$infusion,df.div.exp$temperature)

df.div.exp$type <- sub("A.albopictus","Mosquitoes",df.div.exp$type)
df.div.exp$type <- sub("Microbial Water","Meso. water",df.div.exp$type)

df.div.exp$type <- factor(df.div.exp$type,levels=c("Mosquitoes","Meso. water"))
```

### Richness

```{r}
df.div.exp.se <- summarySE(df.div.exp,measurevar="Observed",groupvars=c("inf.temp","infusion","temperature","type"))

##dots & error bars
gg.rich <- ggplot(df.div.exp,aes(x=infusion,y=Observed,fill=inf.temp,shape=inf.temp))+
  geom_jitter(alpha=0.5,position=position_jitterdodge(jitter.width=0.2),color="gray")+
  geom_errorbar(data=df.div.exp.se,aes(ymax=Observed+se,ymin=Observed-se),width=0.2,color="black",position=position_dodge(width=0.6))+
  geom_point(data=df.div.exp.se,size=2.5,position=position_dodge(width=0.6))+
  #geom_boxplot(outlier.shape=NA,alpha=0.5)+
  ylim(4,34)+
  facet_wrap(~type)+
  theme_cowplot()+
  ylab("ASV richness")+
  xlab("Infusion")+
  scale_x_discrete(labels=c("OL","SG","PW"))+
  scale_fill_manual(values=c("#B5D69F","#7DBA54","#C488A2","#992559","#B6F6EE","#60eede"),name="Infusion, temp.",labels=c("OL_cool","OL_warm","SG_cool","SG_warm","PW_cool","PW_warm"))+
  scale_shape_manual(values=c(22,24,22,24,22,24),name="Infusion, temp.",labels=c("OL_cool","OL_warm","SG_cool","SG_warm","PW_cool","PW_warm"))
gg.rich

##water by time
df.div.water <- subset(df.div.exp,type=="Meso. water")
gg.mw.rich.time <- ggplot(df.div.water,aes(x=infusion,y=Observed,fill=inf.temp))+
  geom_boxplot()+
  #geom_jitter(alpha=0.5,position=position_jitterdodge(jitter.width=0.2),color="gray")+
  #geom_errorbar(data=df.div.exp.se,aes(ymax=Observed+se,ymin=Observed-se),width=0.2,color="black",position=position_dodge(width=0.6))+
  #geom_point(data=df.div.exp.se,size=2.5,position=position_dodge(width=0.6))+
  #geom_boxplot(outlier.shape=NA,alpha=0.5)+
  facet_wrap(~day)+
  theme_cowplot()+
  ylab("ASV richness")+
  xlab("Infusion")+
  scale_x_discrete(labels=c("OL","SG","PW"))+
  scale_fill_manual(values=c("#B5D69F","#7DBA54","#C488A2","#992559","#B6F6EE","#60eede"),name="Infusion, temp.",labels=c("OL_cool","OL_warm","SG_cool","SG_warm","PW_cool","PW_warm"))+
  ggtitle("Mesocosm water")
gg.mw.rich.time
```

### Simpson's

```{r}
df.div.exp.se <- summarySE(df.div.exp,measurevar="InvSimpson",groupvars=c("inf.temp","infusion","temperature","type"))

##dots & error bars
gg.simp <- ggplot(df.div.exp,aes(x=infusion,y=InvSimpson,fill=inf.temp,shape=inf.temp))+
  geom_jitter(alpha=0.5,position=position_jitterdodge(jitter.width=0.2),color="gray")+
  geom_errorbar(data=df.div.exp.se,aes(ymax=InvSimpson+se,ymin=InvSimpson-se),width=0.2,color="black",position=position_dodge(width=0.6))+
  geom_point(data=df.div.exp.se,size=2.5,position=position_dodge(width=0.6))+
  #geom_boxplot(outlier.shape=NA,alpha=0.5)+
  facet_wrap(~type)+
  theme_cowplot()+
  ylab("Simpson's (inv.)")+
  xlab("Infusion")+
  scale_x_discrete(labels=c("OL","SG","PW"))+
  scale_fill_manual(values=c("#B5D69F","#7DBA54","#C488A2","#992559","#B6F6EE","#60eede"),name="Infusion, temp.",labels=c("OL_cool","OL_warm","SG_cool","SG_warm","PW_cool","PW_warm"))+
  scale_shape_manual(values=c(22,24,22,24,22,24),name="Infusion, temp.",labels=c("OL_cool","OL_warm","SG_cool","SG_warm","PW_cool","PW_warm"))
gg.simp

gg.mw.simp.time <- ggplot(df.div.water,aes(x=infusion,y=InvSimpson,fill=inf.temp))+
  geom_boxplot()+
  #geom_jitter(alpha=0.5,position=position_jitterdodge(jitter.width=0.2),color="gray")+
  #geom_errorbar(data=df.div.exp.se,aes(ymax=Observed+se,ymin=Observed-se),width=0.2,color="black",position=position_dodge(width=0.6))+
  #geom_point(data=df.div.exp.se,size=2.5,position=position_dodge(width=0.6))+
  #geom_boxplot(outlier.shape=NA,alpha=0.5)+
  facet_wrap(~day)+
  theme_cowplot()+
  ylab("Simpson's index (inv.)")+
  xlab("Infusion")+
  scale_x_discrete(labels=c("OL","SG","PW"))+
  scale_fill_manual(values=c("#B5D69F","#7DBA54","#C488A2","#992559","#B6F6EE","#60eede"),name="Infusion, temp.",labels=c("OL_cool","OL_warm","SG_cool","SG_warm","PW_cool","PW_warm"))+
  ggtitle("Mesocosm water")
```

### Multi-panel (Figure 2)

```{r}
gg.div <- ggarrange(gg.rich,gg.simp,common.legend=T,legend="right",labels=c("(a)","(b)"))
gg.div

##just water things
ggarrange(gg.mw.rich.time,gg.mw.simp.time,common.legend=T,legend="right",nrow=2)

#ggsave("div.water.time.png",width=6,height=6)
```

## Statz

### Richness{.tabset}

#### Water richness stats

```{r}
df.div.mw <- subset(df.div.exp,type=="Meso. water")

df.div.mw$day <- as.factor(df.div.mw$day)
str(df.div.mw)

#hist(df.div.mw$Observed)
#shapiro.test(log(df.div.mw$Observed))
shapiro.test(df.div.mw$Observed)

#wrich1<-glmmTMB(Observed~day+temperature+infusion+dispersal+offset(log(lib_size_trim))+(1|mesocosm),family="poisson",data=df.div.mw)
#wrich2<-glmmTMB(Observed~day+temperature+infusion+dispersal+offset(log(lib_size_trim))+(1|mesocosm),family="compois",data=df.div.mw)
wrich3<-glmmTMB(Observed~day+temperature+infusion+dispersal+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mw)

#AICtab(wrich1,wrich2,wrich3)

##full model
wrich1<-glmmTMB(Observed~day+temperature+infusion+dispersal+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mw)
##no dispersal
wrich1nd<-glmmTMB(Observed~day+temperature+infusion+offset(log(lib_size_trim))+(1|mesocosm), data=df.div.mw)
##minus infusion
wrich1ni<-glmmTMB(Observed~day+temperature+dispersal+offset(log(lib_size_trim))+(1|mesocosm), data=df.div.mw)
##minus temperature 
wrich1nt<-glmmTMB(Observed~day+infusion+dispersal+offset(log(lib_size_trim))+(1|mesocosm), data=df.div.mw)
##minus time 
wrich1nnd<-glmmTMB(Observed~temperature+infusion+dispersal+offset(log(lib_size_trim))+(1|mesocosm), data=df.div.mw)

anova(wrich1,wrich1nd) #0.3931
anova(wrich1,wrich1ni) #2.2e-16***
anova(wrich1,wrich1nt) #0.4119
anova(wrich1,wrich1nnd) #0.001874**

Anova(wrich1)
# Response: Observed
# Response: Observed
#                Chisq Df Pr(>Chisq)    
# day          12.9404  2   0.001549 ** 
# temperature   0.6742  1   0.411577    
# infusion    151.4172  2  < 2.2e-16 ***
# dispersal     0.7306  1   0.392704    
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

AICtab(wrich1,wrich1nd,wrich1ni,wrich1nt,wrich1nnd) #no temp best... almost tied with no dispersal, then full

##interactions?
wrich1.int <- glmmTMB(Observed~day*infusion+temperature+dispersal+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mw)
wrich1.int1 <- glmmTMB(Observed~day+infusion*temperature+dispersal+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mw)
wrich1.int2 <- glmmTMB(Observed~day+infusion+temperature*dispersal+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mw)
wrich1.int3 <- glmmTMB(Observed~day*temperature+infusion+dispersal+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mw)
wrich1.int4 <- glmmTMB(Observed~day*dispersal+temperature+infusion+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mw)
wrich1.int5 <- glmmTMB(Observed~day+dispersal*infusion+temperature+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mw)

anova(wrich1,wrich1.int) #2.86e-10***
anova(wrich1,wrich1.int1) #ns
anova(wrich1,wrich1.int2) #ns
anova(wrich1,wrich1.int3) #ns
anova(wrich1,wrich1.int4) #ns
anova(wrich1,wrich1.int5) #ns

Anova(wrich1.int)
# Analysis of Deviance Table (Type II Wald chisquare tests)
# 
# Response: Observed
#                 Chisq Df Pr(>Chisq)    
# day           17.4811  2    0.00016 ***
# infusion     167.8537  2  < 2.2e-16 ***
# temperature    0.7595  1    0.38348    
# dispersal      0.7627  1    0.38248    
# day:infusion  58.9948  4  4.717e-12 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

#tapply(df.div.mw$log(Observed), df.div.mw$infusion, mean)
#tapply(df.div.mw$log(Observed), df.div.mw$day, mean)

wrich1.red <- glmmTMB(Observed~day+infusion+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mw)
wrich1.red.int <- glmmTMB(Observed~day*infusion+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mw)

anova(wrich1.red,wrich1.red.int) #sig 4e-10***

##model checking
shapiro.test(residuals(wrich1.red.int)) #awesome
qqnorm(residuals(wrich1.red.int))
qqline(residuals(wrich1.red.int), col="red")

#plotResiduals(wrich1.int)
#wrich1.red.int.resid <- simulateResiduals(fittedModel = wrich1.red.int, plot = T)

##posthoc things
##just without dispersal because not plotting it
##also not plotting day but it should be in the model somewhere already
wrich1.int.em <- emmeans(wrich1.int,~infusion*temperature)

multcomp::cld(wrich1.int.em)
 # infusion temperature emmean    SE  df lower.CL upper.CL .group
 # SW       C             19.1 0.351 198     18.5     19.8  1    
 # SW       H             19.5 0.353 198     18.8     20.1  1    
 # SG       C             21.2 0.355 198     20.5     21.9   2   
 # SG       H             21.5 0.357 198     20.8     22.2   2   
 # OL       C             24.7 0.351 198     24.0     25.4    3  
 # OL       H             25.0 0.353 198     24.3     25.7    3  
```

#### Water richness stats - rarefied

Same as above but without offset for library size. Not knitting this part, just checking out in R

```{r, eval=F}
df.div.mw <- subset(df.div.exp,type=="Meso. water")

df.div.mw$day <- as.factor(df.div.mw$day)
str(df.div.mw)

#hist(df.div.mw$Observed)
#shapiro.test(log(df.div.mw$Observed))
shapiro.test(df.div.mw$Observed)

#wrich1<-glmmTMB(Observed~day+temperature+infusion+dispersal)+(1|mesocosm),family="poisson",data=df.div.mw)
#wrich2<-glmmTMB(Observed~day+temperature+infusion+dispersal)+(1|mesocosm),family="compois",data=df.div.mw)
wrich3<-glmmTMB(Observed~day+temperature+infusion+dispersal+(1|mesocosm),data=df.div.mw)

#AICtab(wrich1,wrich2,wrich3)

##full model
wrich1<-glmmTMB(Observed~day+temperature+infusion+dispersal+(1|mesocosm),data=df.div.mw)
##no dispersal
wrich1nd<-glmmTMB(Observed~day+temperature+infusion+(1|mesocosm), data=df.div.mw)
##minus infusion
wrich1ni<-glmmTMB(Observed~day+temperature+dispersal+(1|mesocosm), data=df.div.mw)
##minus temperature 
wrich1nt<-glmmTMB(Observed~day+infusion+dispersal+(1|mesocosm), data=df.div.mw)
##minus time 
wrich1nnd<-glmmTMB(Observed~temperature+infusion+dispersal+(1|mesocosm), data=df.div.mw)

anova(wrich1,wrich1nd) #0.466
anova(wrich1,wrich1ni) #2.2e-16***
anova(wrich1,wrich1nt) #0.4789
anova(wrich1,wrich1nnd) #0.006693**

Anova(wrich1)
# Analysis of Deviance Table (Type II Wald chisquare tests)
# 
# Response: Observed
#                Chisq Df Pr(>Chisq)    
# day          10.2547  2   0.005932 ** 
# temperature   0.5020  1   0.478604    
# infusion    133.7700  2  < 2.2e-16 ***
# dispersal     0.5322  1   0.465695    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

AICtab(wrich1,wrich1nd,wrich1ni,wrich1nt,wrich1nnd) #no temp best... almost tied with no dispersal, then full

##interactions?
wrich1.int <- glmmTMB(Observed~day*infusion+temperature+dispersal+(1|mesocosm),data=df.div.mw)
wrich1.int1 <- glmmTMB(Observed~day+infusion*temperature+dispersal+(1|mesocosm),data=df.div.mw)
wrich1.int2 <- glmmTMB(Observed~day+infusion+temperature*dispersal+(1|mesocosm),data=df.div.mw)
wrich1.int3 <- glmmTMB(Observed~day*temperature+infusion+dispersal+(1|mesocosm),data=df.div.mw)
wrich1.int4 <- glmmTMB(Observed~day*dispersal+temperature+infusion+(1|mesocosm),data=df.div.mw)
wrich1.int5 <- glmmTMB(Observed~day+dispersal*infusion+temperature+(1|mesocosm),data=df.div.mw)

anova(wrich1,wrich1.int) #4.35e-11***
anova(wrich1,wrich1.int1) #ns
anova(wrich1,wrich1.int2) #ns
anova(wrich1,wrich1.int3) #ns
anova(wrich1,wrich1.int4) #ns
anova(wrich1,wrich1.int5) #ns

Anova(wrich1.int)
# Analysis of Deviance Table (Type II Wald chisquare tests)
# 
# Response: Observed
#                 Chisq Df Pr(>Chisq)    
# day           13.9847  2  0.0009189 ***
# infusion     154.5860  2  < 2.2e-16 ***
# temperature    0.5795  1  0.4464988    
# dispersal      0.5809  1  0.4459484    
# day:infusion  64.8562  4  2.759e-13 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

#tapply(df.div.mw$log(Observed), df.div.mw$infusion, mean)
#tapply(df.div.mw$log(Observed), df.div.mw$day, mean)

wrich1.red <- glmmTMB(Observed~day+infusion+(1|mesocosm),data=df.div.mw)
wrich1.red.int <- glmmTMB(Observed~day*infusion+(1|mesocosm),data=df.div.mw)

anova(wrich1.red,wrich1.red.int) #sig 4e-11**

##model checking
shapiro.test(residuals(wrich1.red.int)) #awesome
qqnorm(residuals(wrich1.red.int))
qqline(residuals(wrich1.red.int), col="red")

#plotResiduals(wrich1.int)
#wrich1.red.int.resid <- simulateResiduals(fittedModel = wrich1.red.int, plot = T)

##posthoc things
##just without dispersal because not plotting it
##also not plotting day but it should be in the model somewhere already
wrich1.int.em <- emmeans(wrich1.int,~infusion*temperature)

multcomp::cld(wrich1.int.em)
```

#### Mosquito richness stats

```{r}
df.div.mq <- subset(df.div.exp,type=="Mosquitoes")

hist(df.div.mq$Observed)
shapiro.test(df.div.mq$Observed)

##different stat families
mrich1 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mq)
mrich2 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+offset(log(lib_size_trim))+(1|mesocosm),family="compois",data=df.div.mq)
mrich3 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+offset(log(lib_size_trim))+(1|mesocosm),family="genpois",data=df.div.mq)
mrich4 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+offset(log(lib_size_trim))+(1|mesocosm),family="poisson",data=df.div.mq)
#mrich4 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+offset(log(lib_size_trim))+(1|mesocosm),family="nbinom1",data=df.div.mq)
#mrich5 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+offset(log(lib_size_trim))+(1|mesocosm),family="nbinom2",data=df.div.mq)

AICtab(mrich1,mrich2,mrich3,mrich4) #gaussian again

##full model
mrich1 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mq)
##no dispersal 
mrich1nd <- glmmTMB(Observed~newday+temperature+infusion+sex+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mq)
##minus infusion 
mrich1ni <- glmmTMB(Observed~newday+temperature+dispersal+sex+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mq)
##minus temperature 
mrich1nt <- glmmTMB(Observed~newday+infusion+dispersal+sex+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mq)
##minus time 
mrich1nnd <- glmmTMB(Observed~temperature+infusion+dispersal+sex+offset(log(lib_size_trim))+(1|mesocosm), data=df.div.mq)
##minus sex
mrich1ns <- glmmTMB(Observed~newday+temperature+infusion+dispersal+offset(log(lib_size_trim))+(1|mesocosm),data=df.div.mq)

Anova(mrich1)

shapiro.test(residuals(mrich1))
qqnorm(residuals(mrich1))
qqline(residuals(mrich1), col="red")

mrich1.resid <- simulateResiduals(fittedModel = mrich1, plot = T) 

##posthoc things
mrich1.em <- emmeans(mrich1,~infusion*temperature)
multcomp::cld(mrich1.em)
##no differences
```

#### Mosquito richness stats - rarefied

Same as above but without offset for library size

```{r, eval=F}
df.div.mq <- subset(df.div.exp,type=="Mosquitoes")

hist(df.div.mq$Observed)
shapiro.test(df.div.mq$Observed)

##different stat families
mrich1 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+(1|mesocosm),data=df.div.mq)
mrich2 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+(1|mesocosm),family="compois",data=df.div.mq)
mrich3 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+(1|mesocosm),family="genpois",data=df.div.mq)
mrich4 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+(1|mesocosm),family="poisson",data=df.div.mq)
#mrich4 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+(1|mesocosm),family="nbinom1",data=df.div.mq)
#mrich5 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+(1|mesocosm),family="nbinom2",data=df.div.mq)

AICtab(mrich1,mrich2,mrich3,mrich4) #compois 

##full model
mrich1 <- glmmTMB(Observed~newday+temperature+infusion+dispersal+sex+(1|mesocosm),family="compois",data=df.div.mq)
##no dispersal 
mrich1nd <- glmmTMB(Observed~newday+temperature+infusion+sex+(1|mesocosm),family="compois",data=df.div.mq)
##minus infusion 
mrich1ni <- glmmTMB(Observed~newday+temperature+dispersal+sex+(1|mesocosm),family="compois",data=df.div.mq)
##minus temperature 
mrich1nt <- glmmTMB(Observed~newday+infusion+dispersal+sex+(1|mesocosm),family="compois",data=df.div.mq)
##minus time 
mrich1nnd <- glmmTMB(Observed~temperature+infusion+dispersal+sex+(1|mesocosm),family="compois", data=df.div.mq)
##minus sex
mrich1ns <- glmmTMB(Observed~newday+temperature+infusion+dispersal+(1|mesocosm),family="compois",data=df.div.mq)

Anova(mrich1)

shapiro.test(residuals(mrich1))
qqnorm(residuals(mrich1))
qqline(residuals(mrich1), col="red")

mrich1.resid <- simulateResiduals(fittedModel = mrich1, plot = T) 

##posthoc things
mrich1.em <- emmeans(mrich1,~infusion*temperature)
multcomp::cld(mrich1.em)
##no differences
```

### Simpson's{.tabset}

#### Water Simpson's stats

```{r}
hist(df.div.mw$InvSimpson)
shapiro.test(df.div.mw$InvSimpson)

hist(log(df.div.mw$InvSimpson))
shapiro.test(log(df.div.mw$InvSimpson))

##normalizing because residuals bad below
bestNormalize(df.div.mw$InvSimpson) #orderNorm
df.div.mw$simp.norm <- bestNormalize(df.div.mw$InvSimpson)$x.t
shapiro.test(df.div.mw$simp.norm)

##just replaced InvSimpson with simp.norm
##full model
wsimp1 <- glmmTMB(simp.norm~day+temperature+infusion+dispersal+(1|mesocosm),data=df.div.mw)
##no dispersal
wsimp1nd <- glmmTMB(simp.norm~day+temperature+infusion+(1|mesocosm),data=df.div.mw)
##minus infusion
wsimp1ni <- glmmTMB(simp.norm~day+temperature+dispersal+(1|mesocosm),data=df.div.mw)
##minus temperature 
wsimp1nt <- glmmTMB(simp.norm~day+infusion+dispersal+(1|mesocosm),data=df.div.mw)
##minus time 
wsimp1nnd <- glmmTMB(simp.norm~temperature+infusion+dispersal+(1|mesocosm),data=df.div.mw)

Anova(wsimp1)
# Analysis of Deviance Table (Type II Wald chisquare tests)
# 
# Response: simp.norm
#                Chisq Df Pr(>Chisq)    
# day          49.7901  2  1.542e-11 ***
# temperature  41.3001  1  1.306e-10 ***
# infusion    229.1055  2  < 2.2e-16 ***
# dispersal     0.0091  1      0.924    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
##~same results with normalized results
##same results with rarefied

AICtab(wsimp1,wsimp1nd,wsimp1ni,wsimp1nt,wsimp1nnd)
##no dispersal good, then full model

##interaction?
wsimp1.int1 <- glmmTMB(simp.norm~day*infusion*temperature+(1|mesocosm), data=df.div.mw)
wsimp1.int2 <- glmmTMB(simp.norm~day*infusion+temperature+(1|mesocosm), data=df.div.mw) #best AIC
wsimp1.int3 <- glmmTMB(simp.norm~day+infusion*temperature+(1|mesocosm), data=df.div.mw)
wsimp1.int4 <- glmmTMB(simp.norm~day*temperature+infusion+(1|mesocosm), data=df.div.mw)

AICtab(wsimp1.int1,wsimp1.int2,wsimp1.int3,wsimp1.int4,wsimp1nd) #int better
#interaction with day & infusion

Anova(wsimp1.int2) #all 0.001 sig level, same with rare
# Response: simp.norm
#                Chisq Df Pr(>Chisq)    
# day           70.733  2  4.370e-16 ***
# infusion     227.491  2  < 2.2e-16 ***
# temperature   39.606  1  3.107e-10 ***
# day:infusion  58.541  4  5.874e-12 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

##all of the below was bad before normalizing
shapiro.test(residuals(wsimp1))
qqnorm(residuals(wsimp1))
qqline(residuals(wsimp1), col="red")

wsimp1.resid <- simulateResiduals(fittedModel = wsimp1, plot = T) 

##posthoc things
wsimp1.em <- emmeans(wsimp1,~infusion*temperature)
multcomp::cld(wsimp1.em)
 # infusion temperature emmean     SE  df lower.CL upper.CL .group
 # SG       C           -1.255 0.0934 202   -1.439  -1.0708  1    
 # SG       H           -0.656 0.0939 202   -0.841  -0.4712   2   
 # OL       C           -0.104 0.0927 202   -0.287   0.0786    3  
 # SW       C            0.442 0.0927 202    0.259   0.6244     4 
 # OL       H            0.495 0.0932 202    0.311   0.6783     4 
 # SW       H            1.040 0.0932 202    0.857   1.2240      5
```

#### Mosquito Simpson's

From manuscript pre-revisions: "In contrast, the Simpson’s index, which integrates evenness into a measure of alpha diversity, of the adult mosquito microbiome was lower in male mosquitoes relative to female mosquitoes (p<0.0001). This trend was associated with the increased dominance of wAlbB in males. The Simpson’s index of the adult mosquito microbiome did not vary significantly with the dispersal, aquatic chemistry, temperature, or time of emergence."

```{r}
hist(df.div.mq$InvSimpson)

df.div.mq$simp.norm <- bestNormalize(df.div.mq$InvSimpson)$x.t #orderNorm
hist(df.div.mq$simp.norm)

##full model
msimp.all <- glmmTMB(simp.norm~newday+temperature+infusion+dispersal+sex+(1|mesocosm), data=df.div.mq)
##without sex
msimp.nosex <- glmmTMB(simp.norm~day+temperature+infusion+dispersal+(1|mesocosm), data=df.div.mq)
##without dispersal
msimp.nodis <- glmmTMB(simp.norm~newday+temperature+infusion+sex+(1|mesocosm), data=df.div.mq)
##without infusion
msimp.noinf <- glmmTMB(simp.norm~newday+temperature+dispersal+sex+(1|mesocosm), data=df.div.mq)
##without temperature
msimp.notem <- glmmTMB(simp.norm~newday+infusion+dispersal+sex+(1|mesocosm), data=df.div.mq)
##without time
msimp.notim <- glmmTMB(simp.norm~temperature+infusion+dispersal+sex+(1|mesocosm), data=df.div.mq)

Anova(msimp.all)
# Analysis of Deviance Table (Type II Wald chisquare tests)
# 
# Response: simp.norm
#               Chisq Df Pr(>Chisq)    
# newday       2.2911  2    0.31805    
# temperature  0.2987  1    0.58469    
# infusion     2.7011  2    0.25909    
# dispersal    2.8332  1    0.09234 .  
# sex         67.5319  1    < 2e-16 ***

AICtab(msimp.all,msimp.nosex,msimp.nodis,msimp.noinf,msimp.notem,msimp.notim)
#no time & no temp tied for best, then no infusion

##all of the below was bad before normalizing
shapiro.test(residuals(msimp.all))
qqnorm(residuals(msimp.all))
qqline(residuals(msimp.all), col="red")

msimp.all.resid <- simulateResiduals(fittedModel = msimp.all, plot = T) 

##posthoc things
msimp.all.em <- emmeans(msimp.all,~infusion*temperature)
multcomp::cld(msimp.all.em)
#no diffs
```

## Infusion water things

```{r}
df.div.inf <- subset(df.div,type=="Infusion water")

df.div.inf.se <- summarySE(df.div.inf,measurevar="Observed",groupvars=c("infusion"))
df.div.inf.se

# gg.iw.rich <- ggplot(df.div.inf,aes(x=infusion,y=Observed,fill=infusion))+
#   geom_jitter(alpha=0.5,position=position_jitterdodge(jitter.width=0.2),color="gray")+
#   geom_errorbar(data=df.div.inf.se,aes(ymax=Observed+se,ymin=Observed-se),width=0.2,color="black",position=position_dodge(width=0.6))+
#   geom_point(data=df.div.inf.se,size=2.5,position=position_dodge(width=0.6))+
#   #geom_boxplot(outlier.shape=NA,alpha=0.5)+
#   theme_cowplot()+
#   ggtitle("Infusion water")
# gg.iw.rich
```

# Diversity correlations 

## Diversity x day

Doesn't look interesting

```{r}
ggplot(df.div.mq,aes(x=day,y=Observed))+
  geom_point()+
  geom_smooth()

ggplot(df.div.mq,aes(x=day,y=InvSimpson))+
  geom_point()+
  geom_smooth()
```

## Diversity - mesocosm vs. mosquito

```{r}
library(reshape)

meso.corr <- merge(df.div.mq,df.div.mw,by="mesocosm")
ggplot(meso.corr,aes(x=InvSimpson.x,y=InvSimpson.y))+
  geom_point()+
  #facet_wrap(~day.y)+
  geom_smooth()

ggplot(meso.corr,aes(x=Observed.x,y=Observed.y))+
  geom_point()+
  geom_smooth()
```

